\paragraph{EasyCrypt}

WP as a bullet list:

\begin{itemize}
\item EasyCrypt is a proof assistant for reasoning about relational
  properties of probabilistic computations with adversarial code.
\item it mains application is he construction and verification of
  game-based cryptographic proofs.
\item it now has a large corpus of proofs for state-of-the art
  primitives.
\item linked with the Jasmin compiler (written in Coq), it provides a
  way to produce proof of security \& functional correctness of
  high-assurance \& high-speed (world-wide) cryptographic primitives
  implementations.
\item EasyCrypt comes with a standard library that has to be
  reimplemented from scratch. This library includes math theorems
  (e.g. finite algebra, elliptic-curves, general algebra, a bit of
  real analysis) and standard data structures. These results could be
  imported from external libraries already developed in other proof
  assistants.
\item The Jasmin-EasyCrypt link is informal and creates a gap in the
  formal proofs of primitives implementation. Having a Dedukti backend
  for EasyCrypt (along with a Dedukti to Coq translator) would allow
  to formalise that link.
\item EasyCrypt has the ability to use ATP for solving logical
  goals. This translation is informal. Here again, the Dedukti
  platform would provide a way to formalise these translations and to
  reduce the TCB of EasyCrypt.
\end{itemize}
\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../propB"
%%% End:
