\begin{enumerate}
\item Express 2-Level Type Theory (2LTT) as an object theory in Dedukti: 2LTT is
  a stepping stone towards encoding more expressive variants of Homotopy Type
  Theory, and it makes explicit proof steps that are omitted in standard type
  theories. Our main challenge will be to design a translation in which as many
  small proof steps of 2LTT as possible can be checked automatically via rewrite
  rules in Dedukti. For steps that cannot be automated in this way, the
  translation needs to elaborate proofs that are equivalent (in an appropriate
  sense) to the original proof steps.
\item Express a core Cubical Type Theory in 2LTT: here, the main challenge is
  that equality in cubical type theory is very expressive. The associated
  coherence condition are unlikely to translate to rewrite rules of Dedukti, and
  we will have to come up with appropriate proof elaborations.
\item Define structures such as cartesian cubical type theory on top of the
  core, and compare these structures.
\item Import the UniMath library into Dedukti and translate it into Cubical Type Theory. The UniMath library has been built upon a modified version of Martin-Löf type theory, extended by the univalence axiom. Its foundations have purposefully been kept simple, to facilitate a translation to a more modern version of Homotopy Type Theory, such as Cubical Type Theory. We will import UniMath into Dedukti, and testdrive our specification of Cubical Type Theory by exporting the UniMath library to it.
\end{enumerate}
