Building an infrastructure like Logipedia requires that most of the
major proof systems contribute to the project. We currently have eighteen out of
the twenty-four systems presented in Figure 1, and fourteen out of the
sixteen European ones.  The only reason why we do not have them all is
because of budget constraints.

This explains why the consortium has to be large.  We have twenty-eight
partners in the consortium which is almost all the groups working on
formal proof technology in Europe.  A smaller consortium would not
reach the universality which is at the heart of the project.  Being such a
large consortium is a strategic decision.

Each partner brings a different expertise to the project.  Eighteen
partners provide their expertise for one specific system.  Eight bring
their expertise on different aspects of automated theorem
proving. Four bring their expertise of some specific library. Eleven,
bring their expertise on a transversal issue: access (two partners),
proof engineering (nine partners), and publishing (one partner).

Finally, there are seven industry partners among the twenty-eight partners.
One is an industrial research center and  six are enterprises.
Three are transversal and five are
focused on one industrial sector: transportation, health care, energy,
security, and education. These six enterprises are small and
medium-sized enterprises. In contrast the enterprises in the club of
industrial users are of very different sizes, from small and medium-sized
enterprises to very large enterprises such as Alstom or Siemens.

\begin{longtable}{|p{0.3\textwidth}|p{0.15\textwidth}|p{0.1\textwidth}|p{0.1\textwidth}|p{0.1\textwidth}|p{0.1\textwidth}|}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hline
{\bf Partner}
&
{\bf Systems}
&
{\bf ATP}
&
{\bf Libraries}
&
{\bf Transverse}
&
{\bf Industrial sector}
\\
\hline
Institut National de Recherche en Informatique et Automatique
&
HOL Light, PVS, Coq, \tlaplus, HoTT
&
&
Flyspeck
&
Proof engineering
&
\\
\hline
Université de Strasbourg
&
&
&
GeoCoq
&
Proof engineering
&
\\
\hline
Institut National Polytechnique de Toulouse
&
Rodin
&
&
&
Proof engineering
&
\\
\hline
Universität Innsbruck
&
Mizar
&
Using FOL proofs
&
&
Proof engineering
&
\\
\hline
Université de Liège
&
&
SMT reasoners and proofs
&
&
&
\\
\hline
Alma Mater Studiorum --- Università di Bologna
&
Matita, Coq
&
&
&
Proof engineering
&
\\
\hline
Matematički fakultet, Univerzitet u Beogradu
&
&
Coherent logic
&
&
Proof engineering
&
\\
\hline
Technische Universität München
&
Isabelle/HOL
&
&
AFP, Probability / analysis
&
&
\\
\hline
Technische Universiteit Delft
&
Agda
&
&
&
&
\\
\hline
Université Paris-Saclay
&
&
Using SMT proofs
&
&
Proof engineering
&
\\
\hline
Friedrich-Alexander Universität Erlangen-Nürnberg
&
&
&
&
Access, Proof engineering
&\\
\hline
University of Leeds
&
HoTT, Coq
&
&
&
Proof engineering
&
\\
\hline
Göteborgs Universitet
&
Agda
&
&
&
&
\\
\hline
Chalmers Tekniska Högskola
&
HOL4
&
&
CakeML
&
&
\\
\hline
Ludwig-Maximilians-Universität München
&
Minlog
&
&
&
&
\\
\hline
Institut Mines-Télécom
&
Atelier B, FoCaLiZe, Coq
&
ATP proofs for Dedukti
&
&
Proof engineering
&
\\
\hline
Uniwersytet w Białymstoku
&
Mizar
&&
MML
&&\\
\hline
OCamlPro
&
&
SMT
&
&
&
\\
\hline
ClearSy
&
Atelier B
&
&
&
&
Trans- portation, Energy
\\
\hline
University of Birmingham
&
HoTT
&
&
UniMath
&
&
\\
\hline
Commissariat à l’Energie Atomique et aux Energies Alternatives
&
Why3, Frama-C
&
ATP cooperation
&
&
&
Energy, transportation, security
\\
\hline
Duale Hochschule Baden-Württemberg
&
&
FOL reasoners, FOL proofs
&
&
&
\\
\hline
Edukera
&
&
&
&
Access
&
Education
\\
\hline
MED-EL Elektromedizinische Geraete GmbH
&
&
&
&
&
Health care
\\
\hline
Prove \& Run
&
ProvenTools
&
&
&
&
Security
\\
\hline
Konrad-Zuse-Zentrum für Informationstechnik Berlin
&
&
&
&
Publishing
&
\\
\hline
Universitatea Alexandru Ioan Cuza din Iasi
&
K Prover
&
&
&
&
\\
\hline
Runtime Verification SRL
&
K Prover
&
&
&
&
\\
\hline
\end{longtable}

%%% Local Variables:
%%%   mode: latex
%%%   mode: flyspell
%%%   ispell-local-dictionary: "english"
%%% End:
