\chapter{Excellence}

\input{intro}

\section{Objectives}

\begin{todo}{}\color{red}
  Describe the overall and specific objectives for the project1, which should be clear, measurable, realistic and achievable within the duration of the project. Objectives should be consistent with the expected exploitation and impact of the project (see section 2).
\end{todo}

\input{objectives}

\section{Relation to the work program}

\begin{todo}{}\color{red}
  Indicate the work programme topic to which your proposal relates, and explain how your proposal addresses the specific challenge and scope of that topic, as set out in the work programme.
\end{todo}

\input{relationworkprogram}

\section{Concept and methodology}

\subsection{Concept}\label{concept}

\begin{todo}{Concept}\color{red}
Describe and explain the overall concept underpinning the project. Describe the main ideas, models or assumptions involved. Identify any inter-disciplinary considerations and, where relevant, use of stakeholder knowledge. Where relevant, include measures taken for public/societal engagement on issues related to the project;
  
Describe how the participating Research Infrastructures will be integrated to provide, at European and, where relevant, global level, an overarching Research Infrastructure service, contributing to structuring the European Research Area, and, where relevant, global cooperation on Research Infrastructures. Describe also how this ensemble will ensure complementarity and coherence with the existing European Research Infrastructures landscape, including ESFRI infrastructures and existing integrating activities for advanced communities;

Describe any national or international research and innovation activities which will be linked with the project, especially where the outputs from these will feed into the project;
\end{todo}

\input{concept}

\subsection{Methodology}

\begin{todo}{Methodology}\color{red}
  Describe and explain the overall methodology, distinguishing, as appropriate, activities indicated in the relevant section of the work programme. Describe in particular how the various types of activities (i.e. networking activities, trans-national and/or virtual access activities, and joint research activities) will form a coherent and coordinated set of components contributing to the same objectives.

  Where relevant, describe how the gender dimension, i.e. sex and/or gender analysis is taken into account in the project’s content.
  Please note that this question does not refer to gender balance in the teams in charge of carrying out the project but to the content of the planned research and innovation activities . Sex and gender analysis refers to biological characteristics and social/cultural factors respectively. For guidance on methods of sex / gender analysis and the issues to be taken into account, please refer to \url{http://ec.europa.eu/research/swafs/gendered-innovations/index_en.cfm?pg=home} 
\end{todo}

\input{methodology}

\section{Ambition}

\begin{todo}{}\color{red}
  Describe the advance your proposal would provide beyond the state-of-the-art, in particular the progress beyond current achievements in terms of integration and services, and the extent the proposed work is ambitious. In details:

  The extent to which the Networking Activities will foster a culture of co-operation between the participants and other relevant stakeholders.

  The extent to which the Access Activities (Trans-national Access and/or Virtual activities) will offer access to state-of-the-art infrastructures, high quality services, and will enable users to conduct excellent research.

  The extent to which the Joint Research Activities will contribute to quantitative and qualitative improvements of the services provided by the infrastructures.

  Describe the innovation potential {\bf (e.g. ground-breaking objectives, novel concepts and approaches, new products, services or business and organisational models)} which the proposal represents. Where relevant, refer to products and services already available on the market. Please refer to the results of any patent search carried out.
\end{todo}

\input{ambition}

\input{relatedprojects}





%%% Local Variables:
%%% mode: latex
%%% TeX-master: "propB"
%%% End:
