\chapter{Impact}\label{chap:impact}

%\ednote{Maximum length for the whole of Section 3 –-- ten pages}

\section{Expected impacts}\label{sec:expected-impact}

%\begin{todo}{}\color{red} Please be specific, and provide only
%  information that applies to the proposal and its
%  objectives. Wherever possible, use quantified indicators and
%  targets.
  
%  * Describe how your project will contribute to:
  
%     - each of the expected impacts mentioned in the work programme,
%     - under the relevant topic;
  
%      - where relevant, any substantial impacts not mentioned in the
%      - work programme, that would enhance innovation capacity;
%      - create new market opportunities, strengthen competitiveness
%      - and growth of companies, address issues related to climate
%      - change or the environment, or bring other important benefits
%      - for society
  
%  * Describe any barriers/obstacles, and any framework conditions
%  * (such as regulation, standards, public acceptance, workforce
%  * considerations, financing of follow-up steps, cooperation of
%  * other links in the value chain), that may determine whether and
%  * to what extent the expected impacts will be achieved. (This
%  * should not include any risk factors concerning implementation, as
%  * covered in section 3.2.)
%\end{todo}

\input{expectedimpact}


\input{ker}

\section{Measures to maximise impact}

{\bf (a) Dissemination and exploitation of results}

\begin{todo}{}\color{red}
  See participant portal FAQ on how to address \underline{dissemination and exploitation} in Horizon 2020.

  Provide a draft {\bf plan for the dissemination and exploitation of the project's results}. In the case of Integrating Activities, these are typically the results of the joint research activities to improve the infrastructure services, the enhanced access provision, and/or common standards, protocols etc. resulting from networking activities. Please note that such a draft plan is an \underline{admissibility condition}, unless the work programme topic explicitly states that such a plan is not required. 

Show how the proposed measures will help to achieve the expected impact of the project. 

The plan, should be proportionate to the scale of the project, and should contain measures to be implemented both during and after the end of the project. 

Your plan for the dissemination and exploitation of the project's results is key to maximising their {\bf impact}. This plan should describe, in a concrete and comprehensive manner, the {\bf area} in which you expect to make an impact and {\bf who} are the potential users of your results.  Your plan should also describe {\bf how} you intend to use the appropriate channels of dissemination and interaction with potential users.

Consider the full range of potential users and uses, including research, commercial, investment, social, environmental, policy-making, setting standards, skills and educational training where relevant.

Your plan should give due consideration to the possible {\bf follow-up} of your project, once it is finished. Its exploitation could require additional investments, wider testing or scaling up. Its exploitation could also require other pre-conditions like regulation to be adapted, or value chains to adopt the results, or the public at large being receptive to your results.

Include a business plan where relevant.

As relevant, include information on how the participants will manage the research data generated and/or collected during the project, in particular addressing the following issues:

    - What types of data will the project generate/collect?

    - What standards will be used?
    
    - How will this data be exploited and/or shared/made accessible for verification and re-use? If data cannot be made available, explain why.
    
    - How will this data be curated and preserved?
    
    - How will the costs for data curation and preservation be covered?

Actions under Horizon 2020 participate in the extended ‘Pilot on Open Research Data in Horizon 2020 ('open research data by default'), except if they indicate otherwise ('opt-out'.)1. Once the action has started (not at application stage) those beneficaries which do not opt-out, will need to create a more detailed Data Management Plan for making their data findable, accessible, interoperable and reusable (FAIR).

You will need an appropriate consortium agreement to manage (amongst other things) the ownership and access to key knowledge (IPR, research data etc.). Where relevant, these will allow you, collectively and individually, to pursue market opportunities arising from the project's results.

The appropriate structure of the consortium to support exploitation is addressed in section 3.3.

    - Include information about any open source software used or developed by the project.

    - Outline the strategy {\bf for knowledge management and protection}. Include measures to provide {\bf open access} (free on-line access, such as the ‘green’ or ‘gold’ model) to peer-reviewed scientific publications which might result from the project.

    Open access must be granted to all scientific publications resulting from Horizon 2020 actions (in particular scientific peer reviewed articles). Further guidance on open access is available in the H2020 Online Manual on the Participant Portal. This obligation does not apply to trans-national access users.

    Open access publishing (also called 'gold' open access) means that an article is immediately provided in open access mode by the scientific publisher. The associated costs are usually shifted {\bf away from readers, and instead (for example) to the university or research institute to which the} researcher is affiliated, or to the funding agency supporting the research. Gold open access costs are fully eligible as part of the grant. Note that if the gold route is chosen, a copy of the publication has to be deposited in a repository as well.

    Self-archiving (also called 'green' open access) means that the published article or the final peer-reviewed manuscript is archived by the researcher - or a representative - in an online repository before, after or alongside its publication. Access to this article is often - but not necessarily - delayed (‘embargo period’), as some scientific publishers may wish to recoup their investment by selling subscriptions and charging pay-per-download/view fees during an exclusivity period.
\end{todo}

\subsection{Dissemination and/or Use of Project Results, and Management of Intellectual Property}\label{sec:dissemination}


- Dissemination  (Cofernce, Summer schools...)

- Clubs

- Textbooks for working mathematicians and math students...

- co-avised PhDs on Logipedia, between Academia and Industry

- proof editing tools

- use cases from indutry

- cooperation with certification authorities

- APIs

- Press real




{\bf (b) Communication activities}

\begin{todo}{}\color{red}
  Describe the proposed communication measures for promoting the project and its findings during the period of the grant. Measures should be proportionate to the scale of the project, with clear objectives.  They should be tailored to the needs of different target audiences, including groups beyond the project's own community.

  See participant portal FAQ on how to address communication activities in Horizon 2020.

  For further guidance on communicating EU research and innovation for project participants, please refer to the H2020 Online Manual on the Participant Portal.
\end{todo}

\input{relatedprojects}






%%% Local Variables: 
%%% mode: LaTeX
%%% TeX-master: "propB"
%%% End: 

% LocalWords:  ednote
