\chapter{Implementation}\label{chap:implementation}

{\color{red} No month 0, start at month 1}

\section{Work plan --- Work packages, deliverables}

\input{workplan}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Management structure, milestones and procedures}

\input{management}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Consortium as a whole}\label{sec:consortium}

\input{consortium}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Resources to be committed}\label{sec:resources}


Table for buget

Comments : no computers, pm, etc.

\begin{todo}{}\color{red}
Please make sure the information in this section matches the costs as stated in the budget table in section 3 of the administrative proposal forms, and the number of person months, shown in the detailed work package descriptions.

Please provide the following:

- a table showing number of person months required (table 3.4a)

- a table showing ‘other direct costs’ (table 3.4b) for participants where those  costs exceed 15\% of the personnel costs (according to the budget  table in section 3 of the administrative proposal forms) and participants providing trans-national access under this project and incurring travels and subsistence costs for supporting users' access.

Please note that the distribution of resources between access (TA/VA), JRA, and NA components must be duly justified.
\end{todo}

{\color{red} A table with WP in columns and parners in line and
  ressources in pm}

\subsection{In-kind contributions on the beneficiary's premises}

As declared in the online proposal submission form, the section 3 "budget", the LOGIPEDIA consortium has declared a number of in-kind contributions. That is why the requested EU Contribution (K) of our consortium is a smaller amount than the max EU contribution (J). The community supporting the LOGIPEDIA project is wide but each partner has been thoroughly selected and is necessary for the success of the project. A budget of 5 millions euros is too limited to support all the efforts of the community. The difference between the total eligible costs for the project and the requested will therefore be supported by some partners as an-kind contribution.

Here are the in-kind contributions of the project for each WP in euros:

\begin{itemize}
\item WP1 (TU Delft):
\item WP2 (USaclay):
\item WP3 (TUM):
\item WP4 (SystemX):
\item WP5 (?):
\item WP6 (UIBK):
\item WP7 (ULeeds):
\item WP8 (INRIA):
\item WP9 (INRIA):
\end{itemize}

\textbf{Florian Rabe (1/03, 10.03)}\\
Our WP has no in-kind contributions.

\textbf{Benedikt Ahrens (09/03, 11.57pm)}\\
contributing 3 PM in-kind to WP8

\textbf{Jesper Cockx (09/03, 11.27)}\\
For the contributions "in kind", we have two partners that fund a certain percentage of the costs themselves:
\begin{itemize}
\item Gothenburg University: 7700 €
\item Chalmers: 9 000 €
\end{itemize}

\textbf{Helmut Schwichtenberg (09/03, 13.33)}\\
This is to let you know that in case the project is selected Kenji Miyamoto and myself will contribute 6 PMs of in-kind contributions for task 6.3 on Minlog.

\textbf{Stephan Merz (09/03, 11.07)}\\
Concerning in-kind contributions, I left comments in the LaTeX sources for section 3, and I will request the partners to send the email that you ask for.

\textbf{Nicola Gambino (09/03, 12.21)}
\begin{itemize}
\item Nicola Gambino will contribute 3 PM to task 6.1 on homotopy / cubical type theory,
\item Nicola Gambino will contribute 3 PM to task 7.1 on proof engineering
\item Michael Rathjen will contribute 1 PM to task 7.1 on proof engineering
\item Paul Shafer will contribute 1 PM to task 7.1 on proof engineering
\end{itemize}

\textbf{Magnus Myreen (09/03, 11.02)}
\begin{itemize}
\item WP3, the only in-kind part is: T4 The Flyspeck library; Sites: Inria
\item This task (WP3 T4) is 100\% in-kind.
\end{itemize}

\textbf{Pascal Fontaine (08/03, 4.02)}
\begin{itemize}
\item On peut ajouter 2 mois "in kind" pour Stuttgart (Stephan Schulz).  1 mois sur T2.1, 1 mois sur T2.2. Pour rappel, Stu a 0 mois financés.
\item On peut ajouter 6 mois "in kind" pour Liège (Pascal Fontaine).  3 mois sur T2.1, 2 mois sur T2.2, 1 mois sur T2.1.  Pour rappel, Lie a 48 mois financés.
\end{itemize}

\subsection{Travel costs and consumables}\label{sec:travel-costs}

\subsection{Subcontracting}\label{sec:subcontracting}

\begin{todo}{from the proposal template}
  If any part of the work is to be sub-contracted by the participant responsible for it,
  describe the work involved and explain why a sub-contract approach has been chosen for
  it.
\end{todo}

The tasks \taskref{instrumentation}{isabelle},
\taskref{libraries}{afp} (both handled by \site{Tum}) and part of
task~\taskref{structuring}{strontorepml} (handled by \site{Fau}) will
be carried out by subcontracting Dr.\ M.\ Wenzel.  Each subcontract
will cover roughly the equivalent of $3$ person-months.  Concretely,
it concerns the export of proof terms and other data from the Isabelle
system.  Wenzel is the main Isabelle developer and has spent the
last $10$ years building the technological prerequisites for the
required work and is the natural person to carry it out.  However, he
has left academia and started his own enterprise that specializes on
Isabelle kernel development and routinely carries out subcontracts for
Isabelle-related research projects.  Therefore, a subcontract is the
best option as developing the necessary expertise in-house would take
an additional 6-12 person-months per task.  \site{Fau} has already worked with
Wenzel in similar subcontracts twice before (including the
OpenDreamKit EU infrastructure project), and the collaborations have
been very effective and efficient. Wenzel obtained his Ph.D. at
\site{Tum} advised by Nipkow.

The task \taskref{instrumentation}{isabelle} (handled by \site{Tum})
requires special assistance by Dr.\ David Matthews (PROLINGUA LTD,
Edinburgh): As provider of the underlying Poly/ML infrastructure,
Matthews is in a unique position to provide extra scalability of ML
heap management, and thus allow Isabelle to export more library
material.

\subsection{Subcontracting costs}\label{sec:subcontracting-costs}

As explained in Section~\ref{sec:subcontracting}, \site{Tum} asks for
EUR~50.000 and \site{Fau} for EUR~20.000 to subcontract Dr.\ M.\ Wenzel to
carry out (part of) the work in task
\taskref{instrumentation}{isabelle}, \taskref{libraries}{afp} and
\taskref{structuring}{strontorepml}. Further EUR~15.000 are required
by \site{Tum} to subcontract Dr.\ D.\ Matthews to participate in
critical parts of task \taskref{instrumentation}{isabelle}.

\subsection{Other costs}

%%% Local Variables: 
%%% mode: LaTeX
%%% TeX-master: "propB"
%%% mode: flyspell
%%% ispell-local-dictionary: "english"
%%% End: 

% LocalWords:  pn newpage site-FAU site-efo site-baz jointpub efo baz
% LocalWords:  jointproj coherencetable assoc-partner
