A detailed {\em Dissemination and communication plan} will be
delivered at month 4 (to be updated in the course of the project as
necessary). That plan will outline all of the activities to be
undertaken, including the identification of partners' participation in
international, European and national conferences and events, relevant
publication channels and opportunities for collaboration and
cooperation with other projects and initiatives. Special attention
will be given to the exploitation strategy to align the dissemination
objectives with the exploitation goals. The Plan will also:
\begin{compactitem}
\item increase the awareness of Logipedia in academia, 
\item use Logipedia as a way to increase the cooperation between academia and
industry,
\item prepare the sustainability and exploitation of Logipedia before the
  end of the project, illustrating the philosophy of 
the 
\href{http://roadmap2018.esfri.eu/media/1048/rm2018-part1-20.pdf}{2018
  roadmap} of the European Strategy Forum on Research Infrastructures
(ESFRI): 
``A robust long-term vision is essential to successfully and
sustainably develop, construct and operate Research Infrastructures.''
even if Logipedia is still in its ``incubation phase''.
\end{compactitem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{(a) Plan for dissemination and exploitation of results}
\label{sec:dissemination}


\subsubsection*{Dissemination}

Logipedia will strictly follow the open access policy of Horizon 2020
by providing online access to scientific information that is free of
charge to the end-user and that is reusable.  The project will combine
different measures to foster open access to knowledge. With reference
to Open Data, Logipedia will comply with the requirements of the Open
Research Data Pilot fully.

Dissemination of information and knowledge will be achieve through
the following means (described in work package 8):

\begin{itemize}
\item   
The participation to conferences. 

In computer science, publishing in conference
proceedings are often favoured over journal publication. We have targeted
more than ten conferences:
\begin{compactitem}
\item
  CADE (Conference on Automated Deduction)
\item
  CICM (Intelligent Computer Mathematics)
\item
  CPP (Certified Programs and Proofs)
\item
  CSL (Computer Science Logic)
\item
  FSCD (Formal Structures for Computation and Deduction)
\item
  FROCOS (Frontiers of Combining Systems)
\item
  ICALP (International Colloquium on Automata, Languages, and Programming)
\item
  IJCAR (International Joint Conference on Automated Reasoning)
\item
  ITP (Interactive Theorem Proving)
\item
LFMTP (Logical Frameworks and Meta-Languages: Theory
and Practice)
\item
  LICS (Logic in Computer Science)
\item
  LPAR (Logic Programming and Automated Reasoning)
\item
  PxTP (Proof eXchange for Theorem Proving)
\end{compactitem}
  
\item Open Access publications.

\item
The organisation of our own yearly event, with a conference, parallel
specialised workshops, and the general assembly.

The organisation of a final event to disseminate and transfer the
knowledge generated with the aim to enable others to use and take up
results, thus maximising the impact of the project. The event will
focus on sharing information and gather feedback through a more
interactive and dynamic approach to boost the uptake of the Logipedia
results.

\item
The organisation of two summer schools open to anyone and not only the
partners.  We shall organise several training sessions targeting the
different communities of users: master and PhD students to teach them
the foundations of Logipedia, teachers to help them use interactive
theorem provers at school and university, and to engineers to help
them use formal methods tools in their work. Such training sessions
are key dissemination events that will accompany the growing of the
Logipedia community and contribute to educate a new generation of
researchers, teachers and engineers.

\item The supervisation of PhD students
to educate a new generation of researchers and engineers, 
Some will have academic and industrial co-advisors.

\item 
The co-building the Logipedia strategy
by participating to joint meetings, such as
the clubs and advisory board.


\item 
The dissemination of Logipedia in in research, in industry, in
education, and in publishing.  The four clubs contribute to the
dissemination of Logipedia in their own ecosystem, by organising
talks, courses, meetings.

\item A discussion with certification authorities about the use of a
  common language across the European Union.

\item The co-production of textbooks with the members of the club of
  users in education.

\item The increase of reproducibility in science by referencing formal
  proofs in a single place.

\item The use of a free licence for data and software to make 
the data is findable, accessible, interoperable and reusable
and develop Open data / Open science / Open innovation.
\end{itemize}

\begin{longtable*}{|p{0.35\textwidth}|p{0.22\textwidth}|p{0.25\textwidth}|}
\hline
{\bf Action}
&
{\bf Target}
&
{\bf Indicator and schedule}
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hline
{\bf Participating to conferences}
&
Researchers.
&
10 papers per year.
\\
\hline
{\bf Publishing in Open access venues}
&
Scientists and students.
&
All the publications of the partners are open.
\\
\hline
{\bf Organising of annual Logipedia conferences}
&
Researchers, industrials.
&
100 participants each year. Four conferences total.
\\
\hline
{\bf Organising summer schools and training sessions}
&
Master and PhD students, researchers, engineers.
&
2 Summer schools. 100 participants total.
\\
\hline
    {\bf Supervising PhD students} 
&
PhD students.
&
3 PhD students start each year.
\\
\hline
{\bf Co-building the Logipedia strategy}
&
Researchers, industrials.
&
Participation to the meetings
every year.
\\
\hline
{\bf Disseminating Logipedia in relevant communities}
&
Research, industry,
education, publishing.
&
At least one event organised by each club every year.
\\
\hline
{\bf Initiating a discussion with certification authorities}
&
Certification agencies and the industry. 
&
Two meetings are organised with several European certification agencies.
\\
\hline
{\bf Writing textbooks}
&
Under- graduate and secondary education students
&
At least one textbook during the project.
\\
\hline
{\bf Using Logipedia to increase reproducibility in science}
&
Publishers and researchers.
&
Researchers outside the consortium use Logipedia as a reference.
\\
\hline
{\bf Using a free licence for data and software} 
&
Scientists and innovators.
&
Delivery of Logipedia at month 14.
\\
\hline
\end{longtable*}

\subsubsection*{Exploitation}

The objective of the exploitation activities is to
to make sure Logipedia remains functional with a
healthy community and development roadmap beyond the project.
The partners will conduct the following actions.

\begin{longtable*}{|p{0.30\textwidth}|p{0.30\textwidth}|p{0.30\textwidth}|}
\hline
{\bf Action}
&
{\bf Stakeholders}
&
{\bf Indicator and schedule}
\\
\hline
%{\bf Monitoring the innovation during the course of the project.}
%&
%The steering commitee, the European project manager, the 
%transfer, innovation, and partnership department of Inria Saclay and
%any relevant member of our partners' institution.
%&
%Meetings organized with the transfer and innovation department
%upon request and at least once a year.
%\\
%\hline
{\bf 1. Build an organisation in charge of managing Logipedia
after the end of the project}.
&
The project management team, volunteer members of the consortium,
after consulting the advisory board.
&
The structure is created at month 48 at the latest.
\\
\hline
{\bf 2. Raise funds to manage Logipedia}.
&
The project management team, after consulting the advisory board.
&
Sponsors and fundings have been identified at month 36 at the latest.
\\
\hline   
{\bf 3. Find a server to permanently host Logipedia}.
&
The project management team, volunteer members of the consortium,
after consulting the advisory board.
&
Logipedia is kept functional.
\\
\hline
{\bf 4. Continue developing Logipedia}
&
New developers taking over.
&
New libraries (for instance that of ACL2 or Nuprl) are integrated.
New features are added to Logipedia.
\\
\hline
{\bf 5. Generate new projects, such as ``Logipedia, security, and
certification'', ``Logipedia and automated theorem proving'',
``Logipedia and the B method''...}
&
Special interest groups within Logipedia. 
&
New communities adopt Logipedia.
\\
\hline
\end{longtable*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{(b) Communication activities}

Communication activities aim at raising the awareness about Logipedia
to potential stakeholders that would not be concerned by our
dissemination actions. It will ensure the growth of the Logipedia
ecosystem and be a way to advertise the work achieved by the partners
and the clubs of users. It also serves the purpose of informing the
European citizen of the research findings she has been financially
contributing to.

Communication activities will be closely linked to dissemination
objectives.  Six person-months of an experienced communication
officer, from the Inria Saclay communication team, are dedicate to the
sole task of communication.  She will work in close cooperation with
the dissemination, communication, and exploitation work package
leader and the project management team.

The preparation of a Dissemination and communication plan in line with
the project dissemination and communication strategy will be one of
the first tasks to be addressed in work package 8. The Plan will identify
fundamental elements of the dissemination and communication strategy,
including:
\begin{compactitem}
\item key messages to be conveyed (what),
\item tools and channels used (how),
\item timing of the planned activities (when),
\item geographical level (local, national, European) (where) providing
  a guide for the project's and partners' dissemination activities to
  maximise the impact of Logipedia.
\end{compactitem}

In this project, we also have the ambition to communicate to the
general public, even if, in the past, these communication activities
have been considered as less important than the dissemination towards
the research, industry, and education communities.

\begin{longtable*}{|p{0.30\textwidth}|p{0.30\textwidth}|p{0.30\textwidth}|}
\hline {\bf Action} & {\bf Target audience} & {\bf Indicator and
  schedule} \\
\hline {\bf Promoting the existence of the project}
including defining the project visual identity,
creating a web site,
designing flyers, posters, and videos, and
publishing press releases.
&
Research, industry, and education.
& Website at month 3.
\\

\hline
{\bf Promoting the results and the values of the project} through
outreach actions: publication of articles in popular science
magazines, online videos\footnote{such as {\tt
https://www.facebook.com/TheatreLaReineBlanche/videos/518698965681202},
a one-minute video hosted by le Th\'e\^atre de la Reine Blanche,
presenting in
French the example given at the beginning of this document.}, and live
events such that the European Researchers Night or {\em La Fête de la
  Science}, focusing, as it is our habit.  & General public.  & At
least five partners organise an outreach activity in their country.
\\ \hline
\end{longtable*}

%%% Local Variables:
%%%   mode: latex
%%%   mode: flyspell
%%%   ispell-local-dictionary: "british"
%%% End:
