\chapter{Members of the consortium}

{\color{red} Download the template and check the part 4 complies with it}

{\color{red} Add a section 4.2 with third parties (including subcontracting)

Who are the subcontractors, why we need them best value for money

For each subcontractor a table in the template

In-kind contribution has to be described here:

The partners are so commited to the successful implementation of the
project that they have agreed to the following in-kind contributions}

\section{Individual Participants}
\label{sec:partners}

% organisations in the order of the submission web site
\input{sites/Inria}
\input{sites/Strasbourg}
\input{sites/Toulouse}
\input{sites/Innsbruck}
\input{sites/Liege}
\input{sites/Bologna}
\input{sites/Belgrade}
\input{sites/TUM}
\input{sites/Delft}
\input{sites/Saclay}
\input{sites/FAU}
\input{sites/Leeds}
\input{sites/Goeteborg}
\input{sites/Chalmers}
\input{sites/LMU}
\input{sites/IMT}
\input{sites/Bialystok}
\input{sites/OCamlPro}
\input{sites/ClearSy}
\input{sites/Birmingham}
\input{sites/DHBW}
\input{sites/Edukera}
\input{sites/MED-EL}
\input{sites/ProveRun}
\input{sites/ZIB}
\input{sites/UAIC}
\input{sites/RV}
\input{sites/CEA}

\section{Third parties involved in the project (including use of third party resources)}

\subsection*{Subcontracting}\label{sec:subcontracting}


{\color{red}


  Do not speak about subcontracting or in-kind here. It is in part 4.2}



The tasks \taskref{instrumentation}{isabelle},
\taskref{libraries}{afp} (both handled by \site{Tum}) and part of
task~\taskref{structuring}{strontorepml} (handled by \site{Fau}) will
be carried out by subcontracting Dr.\ M.\ Wenzel.  Each subcontract
will cover roughly the equivalent of $3$ person-months.  Concretely,
it concerns the export of proof terms and other data from the Isabelle
system.  Wenzel is the main Isabelle developer and has spent the
last $10$ years building the technological prerequisites for the
required work and is the natural person to carry it out.  However, he
has left academia and started his own enterprise that specializes on
Isabelle kernel development and routinely carries out subcontracts for
Isabelle-related research projects.  Therefore, a subcontract is the
best option as developing the necessary expertise in-house would take
an additional 6-12 person-months per task.  \site{Fau} has already worked with
Wenzel in similar subcontracts twice before (including the
OpenDreamKit EU infrastructure project), and the collaborations have
been very effective and efficient. Wenzel obtained his Ph.D. at
\site{Tum} advised by Nipkow.

The task \taskref{instrumentation}{isabelle} (handled by \site{Tum})
requires special assistance by Dr.\ David Matthews (PROLINGUA LTD,
Edinburgh): As provider of the underlying Poly/ML infrastructure,
Matthews is in a unique position to provide extra scalability of ML
heap management, and thus allow Isabelle to export more library
material.

{\bf Cost.} \site{Tum} asks for
EUR~50.000 and \site{Fau} for EUR~20.000 to subcontract Dr.\ M.\ Wenzel to
carry out (part of) the work in task
\taskref{instrumentation}{isabelle}, \taskref{libraries}{afp} and
\taskref{structuring}{strontorepml}. Further EUR~15.000 are required
by \site{Tum} to subcontract Dr.\ D.\ Matthews to participate in
critical parts of task \taskref{instrumentation}{isabelle}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "propB"
%%% End:
