To foster the interoperability of proof systems and the sustainability
and the cross-verification of formal proofs, we propose to collect
them in an online encyclopedia, called Logipedia.  For each proof,
Logipedia will indicate in which systems it can be used and, it will
provide a version of this proof in the theory of these systems.

Such a project will not only foster the use of formal proofs in
research in mathematics and computer science, but also in industry, by
allowing cross-verification, sustainability, and interoperability of
formal proofs, and in education, freeing the teaching of formal proof
technology from being bound to one system.

Such a project can only have a worldwide ambition. However, as a
majority of proof systems are developed in Europe, there is a unique
opportunity for Europe to take the lead on such a project and prepare
the grounds for the economic spinoffs from the project benefiting
European industry. That is why the consortium gathers most of the
European actors active on formal proof systems, while also developing
links with non-European partners.

\begin{center}
  
$\bigstar$ $\bigstar$ $\bigstar$

\end{center}

When defining our objectives, we need indicators to measure our
performance. For some objectives, the common notion of Technological
readiness level applies. But to measure the level of integration of an
existing proof system and associated proof libraries in Logipedia, we
have introduced another metric: the {\em Logipedia integration
  level}.

\begin{framed}
  \vspace{-0.5cm}
\begin{center}
{\bf \Large The Logipedia integration levels (LIL)\label{lil}}
\end{center}

\begin{itemize}
\item[LIL 0:] Nothing has been done.

\item[LIL 1:] The theory implemented in the system has been defined in
  the $\lambda\Pi$-calculus modulo theory and in Dedukti.

\item[LIL 2:] The system has been instrumented so some of its proofs
  can be exported and checked in Dedukti.

\item[LIL 3:] A significant part of the library of the system has been
  exported and checked in Dedukti.

\item[LIL 4:] A significant part of the library of the system has
  been made available in Logipedia.

\item[LIL 5:] A tool has been built to analyze the translated Dedukti proofs
  for the system, detect those that can be expressed in a theory
  weaker than that of the system, and translate those proofs into a
  weaker theory.

\item[LIL 6:] All proofs of the system have been exported, translated,
  and made available in Logipedia.
\end{itemize}
\end{framed}

The ultimate goal of Logipedia is to have all the formal proofs
available to mankind in a single encyclopedia.  A first proof of
concept contains a few hundred lemmas, from the Matita library,
expressed in the theory of six different systems: Matita, Coq, Lean,
HOL Light, Isabelle/HOL, and PVS.

In the next four years, we plan to address, with different ambitions,
the libraries of Agda, Atelier B, Coq, FoCaLiZe, HOL Light, HOL4,
Isabelle/HOL, Matching logic, Matita, Minlog, Mizar, PVS, Rodin,
Smart, TLA+, and Why3.

These systems can be roughly divided into two groups.  For those that
currently have a integration level 1 or higher (Matita, HOL Light,
FoCaLize, Coq, Isabelle/HOL, Agda, Atelier B, Rodin, and HOL4)
we have preliminary results, and we plan to bring them to a high level
of integration.  For those that have a current integration level of 0
(Mizar, TLA$^+$, PVS, Minlog, Why3, Smart, and Matching logic) we are
only starting out and our goals within this project are generally less
ambitious.  These two groups of systems correspond to different
objectives, but both are key to the project. The first ones will
constitute Logipedia in four years and the second ones prepare the
long term future of the infrastructure.

Beyond this focus on proof systems, we also plan to integrate some
proofs coming from automated theorem provers, SAT solvers, SMT
solvers, and model checkers, when these proofs have a manageable
size. To do so, we focus on two formats, TSTP and LFSC, that are
already used in the Automated theorem proving community and in the SMT
solvers community to express proofs.

Very large proofs, requiring large libraries of auxiliary results,
have been developed in some proof systems. These large libraries
constitute an excellent benchmark for the methods developed and
contribute to populating Logipedia with a large number of elements.
We focus, in this project, mostly on five libraries: the Isabelle
Archive of formal proofs, the Isabelle probability and analysis
library, the Coq geometry library the flyspeck library, and the Cake
ML library.  The targeted integration level for all these libraries is
LIL 4.

The integration of systems of the first group, the integration of
automated theorem proving systems, and the integration of these large
libraries will provide the content of Logipedia, and collecting all
these formal proofs in a single infrastructure is {\em per se} a
networking activity.

If we now turn to the access to these formal proofs, our objective is
to develop an infrastructure that make them freely accessible through
a web browser. So, by construction, the access will be trans-national
and virtual. But beyond these two objectives of a trans-national and
virtual access, an important effort will be made to make this
encyclopedia accessible to a large community of specialists and
non-specialists.  This requires to develop an ergonomic web interface,
a package distribution system, and a search engine.

To make this infrastructure accessible, we also need to structure its
content into libraries, books, chapters, etc., reaping the benefit of
the structure (modules, qualified names, etc.) of some of the
libraries we start with and to enrich the data with meta-data.

For these two objectives the Logipedia integration levels are not
relevant and we use the common Technology readiness level indicator.

We have already noticed that for some proof systems, the implemented
theory already been analyzed and related to other systems, so we can
immediately start integrating proofs from these systems into
Logipedia. For others (Mizar, TLA$^+$, PVS, Minlog, Why3, Smart, and
Matching logic), more research work is needed. Understanding how the
theories implemented in these systems can be expressed in Logipedia is
one of our joint research objectives.  We also add here Homotopy type
theory that is a theory, but not a system.

Finally, to be able to export formal proofs to systems different from
that in which they have been developed, we need to analyze what axioms
they use, and, when it is possible, transform them so that they do not
use such an axiom anymore.  In addition, each system has its own
definitions of natural numbers, real numbers, etc. that must be
aligned: structural results proved for one definition of real numbers,
for instance, will be transported to any isomorphic structure,
regardless of the way it has been defined. This is our last objective.
Its performance indicator is the number of proofs we are able to
bring from level LIL 4 to level LIL 5.

\newpage

\begin{longtable}{|p{0.2\textwidth}|p{0.4\textwidth}|p{0.32\textwidth}|}
\hline
Objective
&
Activity to achieve this objective
&
Key performance indicator\\
\hline
Integration of proof systems
&
Integrate the librairies of Matita, HOL Light, FoCaLiZe, Coq,
Isabelle/HOL, Agda, Atelier B, Rodin, and HOL4.
&
\vspace*{-0.41cm}

\hspace*{-0.24cm}
\begin{tabular}{p{0.127\textwidth}|p{0.07\textwidth}|p{0.07\textwidth}}
System & current & targeted\\
\hline
Matita & LIL 5 & LIL 6\\
\hline
HOL Light & LIL 3 & LIL 5\\
\hline
FoCaLiZe & LIL 3 & LIL 5\\
\hline
Coq & LIL 3 & LIL 5\\
\hline
Isabelle/HOL & LIL 2 & LIL 5\\
\hline
Agda & LIL 2 & LIL 4\\
\hline
Atelier B & LIL 1 & LIL 5\\
\hline
Rodin & LIL 1 & LIL 3\\
\hline
HOL4 & LIL 1 & LIL 5\\
\end{tabular}
\\
\hline
Integration of automated theorem proving
&
Integrate proofs of coming from automated
theorem provers and SMT solvers.
&
\vspace*{-0.41cm}

\hspace*{-0.24cm}
\begin{tabular}{p{0.127\textwidth}|p{0.07\textwidth}|p{0.07\textwidth}}
Format & current & targeted\\
\hline
TSTP & LIL 1 & LIL 4\\
\hline
LFSC & LIL 0 & LIL 4\\
\end{tabular}
\\
\hline
Integration of libraries
&
Integrate the Isabelle Archive of formal proofs, the Isabelle
probability and analysis library, the Coq geometry library, the
flyspeck library, and the Cake ML library.
&
\vspace*{-0.41cm}

\hspace*{-0.24cm}
\begin{tabular}{p{0.127\textwidth}|p{0.07\textwidth}|p{0.07\textwidth}}
Library & current & targeted\\
\hline
AFP & LIL 2 & LIL 4\\
\hline
Proba / analysis & LIL 2 & LIL 4\\
\hline
GeoCoq & LIL 2 & LIL 4\\
\hline
Flyspeck & LIL 2 & LIL 4\\
\hline
CakeML & LIL 2 & LIL 4\\
\end{tabular}
\\
\hline
Development of the infrastructure
&
Make Logipedia will be freely accessible through a web browser, and a
package distribution system. Develop a search engine for mathematical formulas.
&
TRL 4
\\
\hline
Structuring the encyclopedia
&
Structure the content of Logipedia into libraries, books, chapters,
etc., reaping the benefit of the structure (modules, qualified names,
etc.) of some of the libraries we start with and to enrich the data
with meta-data.
&
TRL 4
\\
\hline
New theories, new systems
&
Express the theories of Mizar, TLA$^+$, PVS, Minlog, Why3, Smart, and
Matching logic, and Homotopy type theory in Dedukti.
&
\vspace*{-0.41cm}

\hspace*{-0.24cm}
\begin{tabular}{p{0.127\textwidth}|p{0.07\textwidth}|p{0.07\textwidth}}
System & current & targeted \\
\hline
Mizar & LIL 0 & LIL 4\\
\hline
TLA$^+$ & LIL 0 & LIL 2\\
\hline
PVS & LIL 0 & LIL 2\\
\hline
Minlog & LIL 0 & LIL 4\\
\hline
Why3 & LIL 0 & LIL 4\\
\hline
Smart & LIL 0 & LIL 3\\
\hline
Matching logic & LIL 0 & LIL 3\\
\hline
Homototy type theory & LIL 0 & LIL 3\\
\end{tabular}
\\
\hline
Proof engineering
&
Develop algorithm to analyze which axioms are used in a proof, to
eliminate axioms, and to align concepts.
&
From LIL 4 to LIL 5
\\ \hline
\end{longtable}

The seven objectives give a first idea of the shape of the project.

They all contribute to building a new formal proof community, focused
on the values of knowledge sharing, safety, security, privacy, open
access, education, etc.

%%% Local Variables:
%%%   mode: latex
%%%   mode: flyspell
%%%   ispell-local-dictionary: "english"
%%% End:
