\begin{longtable}{|p{0.2\textwidth}|p{0.75\textwidth}|}
\hline
{\bf Access to the best research e-infrastructures}
&
Proof systems and automated theorem provers are research
infrastructures. Each proof system comes with its own library, and
these libraries are also part of the research infrastructure.  Currently
these infrastructures are small, distributed and disconnected.\\
&
\hspace{0.4cm} Logipedia aims at co-building a large, central European
infrastructure from smaller ones by an integration effort.  This
integration effort is substantial but doable. It contributes to the
challenge of bringing to European researchers and engineers effective
and convenient access to the best research infrastructures, in order
to foster further advances in knowledge and technology.
\\
&
\hspace{0.4cm} This idea to structure a networking activity around the
construction and the use of an infrastructure is relatively new in
computer science and mathematics. The novelty is that Logipedia is not
an infrastructure made of computers, but an infrastructure made of
data and of algorithms manipulating these data.
\\
\hline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\bf A starting community} & The objectives of Logipedia have never
been supported under FP7 or Horizon 2020 calls.\\
& \hspace{0.4cm} It mobilizes 29 research groups (21 in academia and 8
in industry), which is almost all the groups working on formal proof
technology in Europe.  The success of the project requires such a
large network.  Indeed, according to Metcalfe's law, the effect of a
network is proportional to the square of the number of connected
users. We can postulate that the effect of an infrastructure, such as
Logipedia, is proportional to the product of the number of proofs it
contains and the number of its potential users, each of them being
proportional to the number of theories it supports. So, such a project
needs a critical mass to succeed. This is also why we have developed a
ecosystem surrounding the project, constituted of four clubs of users:
a club of industrial users, a club of academic users, a club of users
in education, and a club of users in publishing.  These clubs already
have some members, but they will grow during the four years of the
project, preparing the long term future of the infrastructure.\\
& \hspace{0.4cm} These groups are located in 11 European countries.
Some of these countries have a large number of participants, some
others fewer, reflecting the diversity of maturity of the research on
formal methods in Europe. This project will contribute to develop the
formal proof
culture in the European countries where it is still inceptive.\\
& \hspace{0.4cm} We should also point out the originality of this
project within the
European strategy of Research infrastructure.\\
& $\blacktriangleright$ A novelty of this project is that it
investigates how infrastructures can be used in mathematics and in
computer science.\\
& $\blacktriangleright$ It investigates how immaterial infrastructure
can be used to structure a research community, just like material ones
do.\\
& $\blacktriangleright$ It is focused on mathematical {\em a priori}
knowledge, while most infrastructures are focused on {\em a
  posteriori} knowledge, issued from measures, observations,
experiments, and field surveys.\\
& $\blacktriangleright$ It also investigates how a common
infrastructure articulates the relation between research and industry.
\\
\hline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\bf Networking activities, trans-national access, joint research activities}
&
Three work packages described in Part 3
are dedicated to networking activities
to collect the data that constitute the infrastructure: the formal proofs.

\hspace{0.4cm}
Logipedia will be publicly and freely accessible online through any
web browser and trough a package management tool, so trans-national
and virtual access are directly provided. Its administration will be
decentralized in various places in Europe with two mirror sites in
Saclay and in M\"unchen. But we go beyond trans-national and virtual
access.  Two work packages described below are dedicated to improving
the quality of the access to the encyclopedia.\\
&
\hspace{0.4cm}
Finally, this project fosters new joint research activities. First,
between the members of the consortium. Then, between the members of
the consortium and other academic partners, in particular those
developing automated theorem proving systems. Next, between the
consortium and the industrial users of proof systems. Finally, between
the industrial users themselves, as using proofs developed by others
will foster joint developments. Two more work packages are dedicated
to these research activities.  Such joint research activities
currently exist in Europe to some extent.  Such a joint project is a
unique opportunity to strengthen the existing ones and create new
ones.
\\
\hline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\bf Towards standardization}
&
This project will be a stepping stone for a possible standardization
of proof languages.
\\
&
\hspace{0.4cm}
Unlike other communities in computer science, the formal proof
community is somewhat skeptical with respect to standardization. And
indeed, it would not make sense to standardize a theory for proof
systems, just like it would make no sense to propose Euclidean
geometry as a standard all geometers should use.  Instead, such a
cooperative effort can lead to the standardization of a logical
framework, that is a language to express theories and proofs in any of
these theories. Proceeding slowly and in a cooperative thus may change
the attitude towards standardization in the formal proof community and
make a standardization process more widely accepted.
\\
\hline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\bf Inter-disciplinarity}
&
This project includes mathematicians, logicians and computer
scientists. Therefore, it is clearly inter-disciplinary.
Yet, Logipedia is inter-disciplinary in an even more fundamental
way. 
\\

&
\hspace{0.4cm}
Proving properties of a piece of software driving a car or
piloting an aircraft require to formalize part of the physical world
in which this piece of software evolves. In the same way proving
properties of simulation software, requires to formalize some
properties of the simulated object.  Thus, just like mathematics and
software are involved in any part of modern science, formal proof will
eventually spread, over time, in all areas of science.\\
&
\hspace{0.4cm}
One major obstacle to the formalization of, for example, simulations
is that they typically depend on large bodies of knowledge in
mathematics and physics.  Different proof systems may be better
suited for different applications, thus there is a natural tendency
for a diversity of proof formats, just like there is a natural
tendency for diversity in the use of programming languages.  But, when
it comes to formalizing a particular physical simulation, one needs to
have all the necessary knowledge accessible from a single infrastructure.
Because inter-disciplinarity is required to formalize physical world
simulations, a unifying language, such as Logipedia, for exchanging
formal proofs between proof systems is a must-have.\\
\hline
\end{longtable}
%%% Local Variables:
%%%   mode: latex
%%%   mode: flyspell
%%%   ispell-local-dictionary: "english"
%%% End:
