\documentclass[12pt]{book}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{color}

\begin{document}

\chapter{Excellence}

{\Huge \color{red} This document is under construction}

Bugs may kill. Would you fly in an autonomous plane fully driven by a
piece of software that has not been formally verified?  Or formally
verified using a technology that Europe does not master? Today, the
trust in critical systems relies on formal methods, in particular
formal proofs, that guaranty the safety of the people using
transportation systems---autonomous car, metros, trains, planes,
etc.---, health systems---robotic surgery, etc.---, energy provided by
nuclear plants, etc. The crucial role of formal proof is highlighted
by several successes, like the correctness proof of the automatic
Paris metro line 14 \cite{metro14} or the detect and avoid system for
unmanned aircraft system developed by Nasa \cite{Munoz16}.

\thispagestyle{empty}

\begin{figure}
\begin{tabular}{ll}
  {\sc \underline{Abella}}~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~&{\sc Acl2}\\
{\sc \underline{Agda}} &  {\sc \underline{HOL Light}}\\
{\sc \underline{Atelier B}} &  {\sc IMPS}\\
{\sc \underline{Coq}}  &  {\sc \underline{Lean}}\\
{\sc \underline{FoCaliZe}}  &  {\sc Nuprl}\\
{\sc \underline{HOL4}}  &  {\sc \underline{PVS}}\\
{\sc \underline{Isabelle}}  &  {\sc \underline{TLA+}}\\
{\sc \underline{Matita}}\\
{\sc \underline{Minlog}}\\
{\sc \underline{Mizar}}\\
{\sc \underline{Rodin}}\\
\end{tabular}
\caption{Some major proof systems. The European ones are in the first column.
  Those addressed in the project are underlined\label{systems}}
\end{figure}

Unfortunately, the development of formal methods is slowed down by the
multiplicity of proof systems and the lack of a common theory used by
these systems. This restricts interoperability, sustainability,
certification, etc.  each small community being centered around one
theory and one system, and often re-doing work done elsewhere.  For
instance, the metro line 14 has been proved correct in {\sc Atelier
B}, while the Nasa detect and avoid system for unmanned aircraft
system has been proved correct in {\sc PVS}.

There are around twenty major proof systems in the world
(Figure \ref{systems}) and making these systems interoperable would avoid work
duplication, reduce the development time, and allow independent
certification.  After three decades dedicated to the developement of
these systems, allowing such a cooperation between these systems is
the next step in the developement of the formal proof technology.

\section{Objectives}

\input{objectives}

\section{Relation to the work program}

\input{relationworkprogram}

\section{Concept and methodology}

\input{conceptmethodology}

\section{Ambition}

\input{ambition}

\chapter{Impact}

\section{Expected Impact}

\input{expectedimpact}

\section{Measures to maximise impact}

\chapter{Implementation}

\section{Work plan - Work packages, deliverables}

\input{workplan}

\section{Management structure, milestones and procedures}

{\color{red} Is it the right place to speak about the assessment /
  self-evaluation of the project}


\input{management}
 
\section{Consortium as a whole}

\section{Resources to be committed}

\chapter{Members of the consortium}

\section{Participants (applicants)}

\input{participants}

\section{Third parties involved in the project (including use of third party resources)}


\chapter{Ethics and Security}

\section{Ethics}

\section{Security}

%Integrating activities

%(i) Networking

%(ii) Transnational access+

%(iii) Joint research activities

%p. 54

\bibliographystyle{abbrv}
\bibliography{main}


\end{document}

%%% Local Variables:
%%%   mode: latex
%%%   mode: flyspell
%%%   ispell-local-dictionary: "english"
%%% End:


