Univerzitet u Beogradu (Belgrade): Predrag Janičić, Filip Marić, Vesna Marinković, Danijela Simić, Sana Stojanović-Đurđević.

Uniwersytet w Białymstoku (Bialystok): Karol Pąk.

Alma Mater Studiorum – Università di Bologna: Claudio Sacerdoti.

Clearsy: David Deharbe, Thierry Lecomte, Ronan Saillard.

Delft: Jesper Cockx.

Friedrich-Alexander-Universität Erlangen-Nürnberg: Michael Kohlhase, Dennis Müller, Florian Rabe.

G\"oteborg University - Chalmers Tekniska H\"ogskola: Andreas Abel, Magnus Myreen, Ulf Norell.

Universität Innsbruck: Joshua Chen, Cezary Kaliszyk, Miroslav Olšák, Stanisław Purgał.

Leeds: Nicola Gambino, Michael Rathjen, Paul Shafer.

Université de Liège: Pascal Fontaine.

Technische Universität München: Tobias Nipkow.

Sketis: Makarius Wenzel.

LMU München: Helmut Schwichtenberg. Kenji Miyamoto 

Inria Nancy – Grand Est: Stephan Merz.

Inria Paris: Émilio Gallego, Hugo Herbelin, Théo Zimmerman.

České vysoké učení technické v Praze (Prague): Thibault Gauthier, Martin Suda, Josef Urban.

Inria Saclay – Île de France: Bruno Barras, Frédéric Blanqui, Valentin Blot, Guillaume Burel, Kaustuv Chaudhuri, Gilles Dowek, Catherine Dubois, Georges Gonthier, Olivier Hermant, Jean-Pierre Jouannaud, Chantal Keller, Dale Miller, Pierre-Yves Strub, Burkhart Wolff.

Inria Sophia Antipolis - Méditerranée: Yves Bertot, Pierre Boutry, Cyril Cohen.

University of Southampton: Michael Butler, Thai Son Hoang, Andrew Sogokon.

Université de Strasbourg: Arthur Charguéraud, Nicolas Magaud, Julien Narboux, Pascal Schreck.

National Polytechnique Institute of Toulouse: Yamine Ait Ameur, Jean-Paul Bodeveix, Mamoun Filali. 


