The proposal meets several of the goals of the call INFRAIA-02-2020:
Integrating activities for starting communities (listed on page 56 of
the part 4 of Horizon 2020 Work Programme 2018-2020). It also clearly
refers to one of the mentioned cross-cutting activities: open science.

\paragraph{Wider and more efficient access --- New research infrastructure}
In a shared public encyclopedia providing virtual access, each user,
in research, industry, and eucation, can find the formal proofs she
needs in the logic she wants, regardless the logic and system this
proof has been developed in.

{\color{red} Where does this sentence should be: 
A data management plan will be provided according to Inria's
compliance policy with EU regulations. We also engage ourselves to
make the evolution of the logipedia e-infrastructure compliant with
the European charter for access to research infrastructures.}

\paragraph{Synergy --- Open research data.}
Instead of having a scattered community, each group developing a
library for its own logic and its own system, researchers and
engineers will be able to work together on common developments,
reusing proofs developed in other systems and in other communities.

%In terms of networking, we have already organized one logipedia
%meeting and the funding will insure we can continue to organize
%large-scale international meetings on a regular basis. The first
%logipedia event has proven to be very valuable in terms of exchange of
%best practices.

\paragraph{Partnership with industry}

Several European and non European companies are member of the project,
as contributors or as members of the future {\sc Logipedia} user's group.

\paragraph{Education --- Closer interaction between a larger number of
researchers.}
Education to formal methods in computer science and to formal proofs
in mathematics always hits the same obstacle: the need to choose a
specific theory or system, in contradiction with the claimed
universality of logical truth. Education to formal methods and formal
proofs will gain in universality once it will be demonstrated that
this choice amounts to include, or not, a few axioms and reduction
rules. We also defend that this renewal of logic education at
university level and before is of prime importance in our ``post-truth
era''.

\paragraph{Better management of the continuous flow of data.}
A shared encyclopedia allows a better sustainability of the formal
proofs developed over time. Too many formal proofs developed in the
past are not available any more.

%%% Local Variables:
%%%   mode: latex
%%%   mode: flyspell
%%%   ispell-local-dictionary: "english"
%%% End:
