\begin{sitedescription}{Del}

\logo{TUDelft}

The Programming Languages Research Group at TU Delft is an
internationally leading research group in programming languages, and
active in areas such as language engineering, language design,
domain-specific languages, software verification, and programme logics.
Specifically, Dr. Jesper Cockx is an expert on the Agda system.

\paragraph{Main tasks:}

\begin{compactitem}
  \item \WPref{instrumentation}, task \taskref{instrumentation}{agda}:
  Encoding of features that rely on type-directed conversion -- such
  as eta-equality and definitional irrelevance -- in Dedukti (which
  does not have type-directed conversion).
  \item \WPref{instrumentation}, task \taskref{instrumentation}{agda}:
  Investigating possible designs for a core language for Agda, in
  order to facilitate the exporting of Agda developments to Logipedia.
  \item \WPref{instrumentation}, task \taskref{instrumentation}{agda}:
  Improving the current state-of-the-art on dependently typed
  languages with user-defined rewrite rules on areas such as
  confluence and termination checking.
\end{compactitem}

\paragraph{Publications, products or services:}

\begin{itemize}
  \item Jesper Cockx, Dominique Devriese, Frank Piessens. ``Pattern
  matching without K.'' Proceedings of the 19th ACM SIGPLAN
  International Conference on Functional programming, ACM, 2014.
  \item Jesper Cockx, Dominique Devriese, Frank Piessens. ``Unifiers
  as equivalences: proof-relevant unification of dependently typed
  data.'' Proceedings of the 21st ACM SIGPLAN International Conference
  on Functional Programming, ACM, 2016.
  \item Jesper Cockx, Andreas Abel. ``Elaborating dependent
  (co)pattern matching.'' Proceedings of the ACM on Programming
  Languages, 2(ICFP), 2018.
  \item Gaëtan Gilbert, Jesper Cockx, Matthieu Sozeau, Nicolas
  Tabareau. ``Definitional proof-irrelevance without K.'' Proceedings
  of the ACM on Programming Languages, 3, 2019.
\end{itemize}

%\keypubs{DBLP:conf/icfp/CockxDP14,DBLP:conf/icfp/CockxDP16,DBLP:journals/pacmpl/CockxA18}

\paragraph{Previous projects or activities:}

\paragraph{Infrastructures or technical equipments:}

\paragraph{Persons primarily responsible for carrying out the proposed activities:}

\begin{itemize}

  \item{\bf Jesper Cockx} (leader of work package \WPref{instrumentation}
  and task \taskref{instrumentation}{agda}) is an expert on the theory and
  implementation of Agda, a dependently typed programming language and
  proof assistant that is widely used within the programming languages
  community and beyond. He has worked on both foundational parts of
  Agda such as elaboration of dependent (co)pattern matching and new
  extensions such as rewrite rules and definitional proof
  irrelevance. He is also one of the main contributors to the
  implementation of Agda.

\end{itemize}

% \paragraph{Specific expertise:}

% \begin{compactitem}
% \item Elaboration of high-level programming techniques such as
%   dependent pattern matching to a low-level representation that can be
%   more easily transferred to other languages.
% \item Extending dependently typed languages with user-defined rewrite
%   rules that can be used to encode the theories of other languages.
% \item Automatic translation of Agda developments to Dedukti
%   (collaboration with Guillaume Genestier from Paris-Saclay).
% \end{compactitem}


\end{sitedescription}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../propB"
%%% End:

% LocalWords:  site-jacu.tex clange sitedescription emph compactitem pn semmath
% LocalWords:  prosuming-flexiformal KohSuc asemf06 GinJucAnc alsaacl09 StaKoh
% LocalWords:  tlcspx10 KohDavGin psewads11 ednote Radboud Bia ystok CALCULEMUS
% LocalWords:  textbf keypubs OntoLangMathSemWeb uwb Deyan Ginev Stamerjohanns
% LocalWords:  searchability
