\begin{sitedescription}{Med}

\logo{MED-EL}

MED-EL is the leading European Cochlear Implant (CI) company with worldwide operations. Since its founders developed one of the world’s first cochlear implants in 1975, MED-EL has set new standards, developing and manufacturing technologically advanced hearing implant solutions. With over 1,500 employees across more than 100 countries, MED-EL is the global leader in ground-breaking advances and superior clinical outcomes. MED-EL has continually driven progress through research initiatives at our state-of-the-art R\&D campus and through partnerships with research institutes, universities, and clinics worldwide.

\paragraph{Main tasks:}

\ednote{its main tasks, with an explanation of how its profile matches the tasks in the proposal}

\ednote{specify the main tasks and reference the respective work packages}

\begin{compactitem}
\item Case study of the application of formal verification in the industrial setting in \WPref{theories}.
\end{compactitem}

\paragraph{Publications, products or services:}

\ednote{a list of up to 5 relevant publications, and/or products, services (including widely-used datasets or software), or other achievements relevant to the  call content}

\begin{itemize}
\item Mark H. Burstein, Christoph Bussler, Michal Zaremba, Timothy W. Finin, Michael N. Huhns, Massimo Paolucci, Amit P. Sheth, Stuart K. Williams:
A Semantic Web Services Architecture. IEEE Internet Computing 9(5): 72-81 (2005)
\item Michal Zaremba, Matthew Moran: Enabling execution of Semantic Web Services - WSMX core platform. WIW 2004
\item Hui Shen, Brian Wall, Michal Zaremba, Yuliu Chen, Jim Browne: Integration of business modelling methods for enterprise information system analysis and user requirements gathering. Computers in Industry 54(3): 307-323 (2004)
\item Charles J. Petrie, Holger Lausen, Michal Zaremba: SWS Challenge - First Year Overview. ICEIS (4) 2007: 407-412
\item Tomas Vitvar, Adrian Mocan, Mick Kerrigan, Michal Zaremba, Maciej Zaremba, Matthew Moran, Emilia Cimpian, Thomas Haselwanter, Dieter Fensel: Semantically-enabled service oriented architecture : concepts, technology and application. Service Oriented Computing and Applications 1(2): 129-154 (2007)
\end{itemize}

\paragraph{Previous projects or activities:}

\ednote{a list of up to 5 relevant previous projects or activities, connected to the subject of this proposal}

MED-EL has previously participated in several EU-funded as NeuEar, HEAR-EU, NANOCI and NANOEAR. Michal Zaremba and Daniel Winkler as employees of other organizations participated in SWS, SUPER, SOA4ALL, COIN, DIP, LarKC, SEALS Knowledge Web and several other EU-funded projects.

\paragraph{Infrastructures or technical equipments:}

\ednote{a description of any significant infrastructure and/or any major items of technical equipment, relevant to the proposed work}

\paragraph{Persons primarily responsible for carrying out the proposed activities:}

\begin{itemize} % in alphabetical order

\item{\bf Dr. Michal Zaremba} is Head of Applied Artificial Intelligence Group and is responsible for the innovation of AI based solutions.

\item{\bf Dr. Daniel Winkler} is Research Scientist at Applied AI Gorup. He is responsible for the development of frontend systems for speech and hearing rehabilitation.
    
\end{itemize}

\end{sitedescription}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../propB"
%%% End:
