\begin{enumerate}
\item Univerzitet u Beogradu (Belgrade): Predrag Janičić, Filip Marić, Vesna Marinković, Danijela Simić, Sana Stojanović-Đurđević.
\item Uniwersytet w Białymstoku (Bialystok): Karol Pąk.
\item Alma Mater Studiorum – Università di Bologna: Claudio Sacerdoti.
\item Clearsy: David Deharbe, Thierry Lecomte, Ronan Saillard.
\item Delft: Jesper Cockx.
\item Friedrich-Alexander-Universität Erlangen-Nürnberg: Michael Kohlhase, Dennis Müller, Florian Rabe.
\item G\"oteborg University - Chalmers Tekniska H\"ogskola: Andreas Abel, Magnus Myreen, Ulf Norell.
\item Universität Innsbruck: Joshua Chen, Cezary Kaliszyk, Miroslav Olšák, Stanisław Purgał.
\item Leeds: Nicola Gambino, Michael Rathjen, Paul Shafer.
\item Université de Liège: Pascal Fontaine.
\item Technische Universität München: Tobias Nipkow, Makarius Wenzel (subcontractor)
\item LMU München: Helmut Schwichtenberg. Kenji Miyamoto 
\item Inria Nancy – Grand Est: Stephan Merz.
\item Inria Paris: Émilio Gallego, Hugo Herbelin, Théo Zimmerman.
\item České vysoké učení technické v Praze (Prague): Thibault Gauthier, Martin Suda, Josef Urban.
\item Inria Saclay – Île de France: Bruno Barras, Frédéric Blanqui, Valentin Blot, Guillaume Burel, Kaustuv Chaudhuri, Gilles Dowek, Catherine Dubois, Georges Gonthier, Olivier Hermant, Jean-Pierre Jouannaud, Chantal Keller, Dale Miller, Pierre-Yves Strub, Burkhart Wolff.
\item Inria Sophia Antipolis - Méditerranée: Yves Bertot, Pierre Boutry, Cyril Cohen.
\item University of Southampton: Michael Butler, Thai Son Hoang, Andrew Sogokon.
\item Université de Strasbourg: Arthur Charguéraud, Nicolas Magaud, Julien Narboux, Pascal Schreck.
\item National Polytechnique Institute of Toulouse: Yamine Ait Ameur, Jean-Paul Bodeveix, Mamoun Filali. 
\end{enumerate}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
